---
title: "Data Visualization with ggplot2 - Part 2"
author: "Paris Zhang"
date: "4/19/2020"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_depth: 4
    toc_float: true
    fig_width: 6
    fig_height: 4
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
```


# 1. Statistics


## 1.1 Stats & geoms

### Stats

Exampe 1. `geom_` <-> `stat_`

* By default, `geom_histogram()` calls `stat_bin()` for histograms, binwidth = range/30.
* See `geom_bar()` shows different bins.

```{r}
p <- ggplot(iris, aes(x = Sepal.Width))
p1 <- p + geom_histogram() + labs(title = "geom_histogram()")
p2 <- p + geom_bar() + labs(title = "geom_bar()")
p3 <- p + stat_bin() + labs(title = "stat_bin()")

grid.arrange(p1,p2,p3, nrow=3)
```





### Geoms


## 1.2 Stats outside geoms




***


# 2. Coordinates and Facets


## 2.1 Coordinates layer


## 2.2 Facets layer




***



# 3. Themes


## 3.1 Themes from scratch


## 3.2 Recycling themes




***



# 4. Best Practices



## 4.1 Avoid dynamite plot (barplot)



## 4.2 Avoid pie charts


## 4.3 Avoid heat maps


***





# 5. Case Study

CHIS - California Health Interview Survey dataset from 2009


## 5.1 CHIS - Descriptive stats


## 5.2 CHIS - Mosaic plots


